Online Multiclass Linear Classification with Bandit Feedback

### 2018-12-01 arxiv-version/main.tex
More work on polynomials

We define several univariate polynomials
\begin{align*}
P_n(z) & = (z - 1) \prod_{i=1}^n (z - 2^i)^2 && \text{for $n \ge 0$,} \\
A_{n,k}(z) & = (P_n(z))^k - (P_n(-z))^k && \text{for $n,k \ge 0$,} \\
B_{n,k}(z) & = - (P_n(z))^k - (P_n(-z))^k && \text{for $n,k \ge 0$.}
\end{align*}
Let $r = 2 \left\lceil \frac{1}{2} \log(4m + 1) \right\rceil + 1$ and $s = \left \lceil \log(1/\gamma) \right \rceil$.
We define the polynomial $p:\R^d \to \R$ as
$$
p(x) = \sum_{i=1}^m A_{s,r}\left( \frac{\ip{v_i}{x}}{\gamma} \right) \prod_{\substack{j ~:~ 1 \le j \le m \\ j \neq i}} B_{s,r} \left( \frac{\ip{v_i}{x}}{\gamma} \right)
- \left(m - \frac{1}{2} \right) \prod_{j=1}^m B_{s,r} \left( \frac{\ip{v_i}{x}}{\gamma} \right) \; .
$$
For convenienence we define univariate rational function
\begin{align*}
S_{n,k}(z) & = \frac{A_{n,k}(z)}{B_{n,k}(z)} && \text{for $n,k \ge 0$,}
\end{align*}
and a multivariate rational function
$$
Q(x) = \left( \sum_{i=1}^m S_{s,r}\left( \frac{\ip{v_i}{x}}{\gamma} \right) \right) - \left(m - \frac{1}{2} \right) \; .
$$
It easy to verify that
$$
p(x) = Q(x) \prod_{i=1}^m B_{s,r} \left( \frac{\ip{v_i}{x}}{\gamma} \right) \; .
$$

\begin{proposition}[Properties of $P_n$]
\hspace{1cm} % Dummy space
\begin{enumerate}
\item If $z \in [0,1]$ then $P_n(-z) \le P_n(z) \le 0$.
\item If $z \in [1,2^n]$ then $0 \le 4P_n(z) \le -P_n(-z)$.
\end{enumerate}
\end{proposition}

\begin{proof}
To prove the first part, note that $P_n(z)$ and $P_n(-z)$ are non-positive for
$z \in [0,1]$. If $z=1$ then $P_n(z)=P_n(-z) = 0$. For $z \in [0,1)$, we have
$$
\frac{P_n(-z)}{P_n(z)} = \frac{1+z}{1-z} \prod_{i=1}^n \frac{(z+2^i)^2}{(z-2^i)^2} \ge 1 \; .
$$
TODO
\end{proof}

\begin{proposition}[Properties of $S_{n,k}$]
\hspace{1cm} % Dummy space
\begin{enumerate}
\item
\end{enumerate}
\end{proposition}

\begin{proof}
TODO
\end{proof}

>>>>

We define several univariate polynomials
\begin{align*}
P_n(z) & = (z - 1) \prod_{i=1}^n (z - 2^i)^2 && \text{for $n \ge 0$,} \\
A_{n,k}(z) & = (P_n(z))^k - (P_n(-z))^k && \text{for $n,k \ge 0$,} \\
B_{n,k}(z) & = - (P_n(z))^k - (P_n(-z))^k && \text{for $n,k \ge 0$.}
\end{align*}
Let $r = 2 \left\lceil \frac{1}{2} \log(4m + 1) \right\rceil + 1$ and $s = \left \lceil \log(1/\gamma) \right \rceil$.
We define the polynomial $p:\R^d \to \R$ as
$$
p(x) = \left[ \sum_{i=1}^m A_{s,r}\left( \frac{\ip{v_i}{x}}{\gamma} \right) \prod_{\substack{j ~:~ 1 \le j \le m \\ j \neq i}} B_{s,r} \left( \frac{\ip{v_j}{x}}{\gamma} \right) \right]
- \left(m - \frac{1}{2} \right) \prod_{j=1}^m B_{s,r} \left( \frac{\ip{v_j}{x}}{\gamma} \right) \; .
$$
For convenienence we define univariate rational function
\begin{align*}
S_{n,k}(z) & = \frac{A_{n,k}(z)}{B_{n,k}(z)} && \text{for $n,k \ge 0$,}
\end{align*}
and a multivariate rational function
$$
Q(x) = \left( \sum_{i=1}^m S_{s,r}\left( \frac{\ip{v_i}{x}}{\gamma} \right) \right) - \left(m - \frac{1}{2} \right) \; .
$$
It easy to verify that
$$
p(x) = Q(x) \prod_{j=1}^m B_{s,r} \left( \frac{\ip{v_j}{x}}{\gamma} \right) \; .
$$

\begin{proposition}[Properties of $P_n$]
\label{proposition:properties-of-p-n}
\hspace{1cm} % Dummy space
\begin{enumerate}
\item If $z \in [0,1]$ then $P_n(-z) \le P_n(z) \le 0$.
\item If $z \in [1,2^n]$ then $0 \le 4P_n(z) \le -P_n(-z)$.
\end{enumerate}
\end{proposition}

\begin{proof}
To prove the first part, note that $P_n(z)$ and $P_n(-z)$ are non-positive for
$z \in [0,1]$. For $z=1$, $P_n(z) = 0$. For $z \in [0,1)$, we have
$$
\frac{P_n(-z)}{P_n(z)} = \frac{1+z}{1-z} \prod_{i=1}^n \frac{(z+2^i)^2}{(z-2^i)^2} \ge 1
$$
since each of the fractions is at least $1$.

To prove the second part, notice that $P_n(z)$ is non-negative and $P_n(-z)$ is
negative for any $z \in [1,2^n]$. Now, fix $z \in [1,2^n]$ and let $j \in
\{1,2,\dots,n\}$ be such that $2^{j-1} \le z \le 2^j$. This implies that
$(z+2^{j})^2 \ge 4 (z - 2^j)^2$. Therefore,
$$
\frac{-P_n(-z)}{P_n(z)}
= \frac{1+z}{z-1} \cdot \frac{(z+2^j)^2}{(z-2^j)} \prod_{\substack{i ~:~ 1 \le i \le n \\ i \neq j}} \frac{(z+2^j)^2}{(z-2^j)}
\ge 4 \;
$$
since one fraction is at least $4$ and all other fractions are at least $1$.
\end{proof}

\begin{proposition}[Properties of $S_{n,k}$]
For any integer $n \ge 0$,
\begin{enumerate}
\item If $z \in [1,2^n]$ then $S_{n,r}(z) \in [1,1+\frac{1}{2m}]$.
\item If $z \in [-2^n, -1]$ then $S_{n,r} \in [-1-\frac{1}{2m}, -1]$.
\item If $z \in [-1,1]$ then $|S_{n,r}(z)| \le 1$.
\item If $z \in [-2,2^n]$ then $B_{n,r}(z) \ge \left(1 - \frac{1}{4m+1} \right) 2^{n(n-1)r}$
\end{enumerate}
\end{proposition}

\begin{proof}
Note that $B_{n,r}(z)$ is an even function and $A_{n,r}(z)$ is an odd function.
Therefore, $S_{n,r}(z)$ is odd. Also notice that $r$ is an odd integer.

\begin{enumerate}
\item Observe that $S_{n,r}(z)$ can be written as
$$
S_{n,r}(z) = \frac{1 + \left( - \frac{P_n(z)}{P_n(-z)}\right)^r}{1 - \left( - \frac{P_n(z)}{P_n(-z)}\right)^r} \; .
$$
Let $c = \left( - \frac{P_n(z)}{P_n(-z)}\right)^r$. Since $z \in [1,2^n]$, by
part 2 of~\autoref{proposition:properties-of-p-n}, $c \in [0,\frac{1}{4^r}] =
[0,\frac{1}{4m+1}]$. Thus, $S_{n,r}(z) = \frac{1+c}{1-c} \in [1,1 +
\frac{1}{2m}]$.

\item The statement follows part 1 and the fact that $S_{n,r}(z)$ is odd.

\item Observe that $S_{n,r}(z)$ can be written as
$$
S_{n,r}(z) = \frac{1 + \left( - \frac{P_n(z)}{P_n(-z)}\right)^r}{1 - \left( - \frac{P_n(z)}{P_n(-z)}\right)^r}
$$
and let $c = \left( - \frac{P_n(z)}{P_(-z)}\right)^r$. If $z \in [0,1]$, by part
1 of~\autoref{proposition:properties-of-p-n} and the fact that $r$ is odd, $c
\in [-1,0]$, and thus, $S_{n,r}(z) = \frac{1+c}{1-c} \in [0,1]$. Since
$S_{n,r}(z)$ is odd, for $z \in [-1,0]$, $S_{n,r}(z) \in [-1,0]$.

\item TODO

\end{enumerate}
\end{proof}

Clearly, $\deg(A_{s,r}) = \deg(B_{s,r}) = (s+1)r$ and thus $\deg(p) = (s+1)rm$.
TODO
